--- 
title: "Installing Torsten"
subtitle: | 
  Using the bbr.bayes `install_torsten()` function
order: 10000
categories: 
- Torsten
tbl-colwidths: [25,20, 45]
toc: true
toc-depth: 2
---


# Introduction

Torsten is a fork of Stan containing additional Stan language functions to facilitate implementation of pharmacometric models. It plays a role analogous to prediction for population pharmacokinetics (PREDPP) for NONMEM® models. The simplest way to install Torsten is to use the `install_torsten()` function from the `bbr.bayes` package.

This page demonstrates how to install Torsten using the `install_torsten()` function.

# Set up

<hr />

Load required packages and set file paths to your model and figure directories. 

```{r}
library(bbr)
library(bbr.bayes)
library(here)

```

# Basic installation

The `install_torsten()` function attempts to download and install the latest release of Torsten. The following statement installs Torsten as a folder named "Torsten" in the project folder. By default, the installation will not overwrite a previous installation of the same Torsten version at that location. You can overwrite a previous installation by adding `overwrite = TRUE`. You may want to do that in cases where you want to change the C++ compiler options via the `cpp_options` argument.

```{r, eval = TRUE}

install_torsten(dir = here("Torsten"))

```

# Enabling multi-threading

One strategy for reducing computation time of Stan/Torsten analyses is to use the `reduce_sum` Stan function to allow within-chain parallel computation via multi-threading. That requires specific compiler options when Torsten is installed. The following statement installs Torsten using the compiler options needed for multi-threading.

```{r, eval = TRUE}

install_torsten(dir = here("Torsten"), overwrite = TRUE,
                cpp_options = list(stan_threads = TRUE))

```


